A Study of Common Bug Fix Patterns in Rust

### 2022-11-28 discussion.tex
linters became more explicit
minor edit

In this work, we used Ruxanne to successfully mine 20 cross-project bug fix patterns. We presented these patterns in two groups: 12 general patterns and 8 BC-related patterns. Clippy is able to detect five of the BC-related patterns. Fundamentally, Clippy cannot detect some of the remaining patterns. For instance, we proposed the pattern \textit{Adding mutability}. In the Rust context, this change anticipates a future change to the code which will perform a mutation of a previously immutable value. Predicting the future is beyond Clippy's remit. To the best of our knowledge, the rest of the patterns cannot be reported by the current linting tools.

In each group, we categorized the bug fix patterns based on their underlying program element. The patterns encompass a wide variety of program elements both in the general and BC-related groups. This shows that our weighting scheme is not biased towards specific program elements. Also, our patterns are all cross-project fix patterns: each fix pattern has been seen at least in three different projects.

@@

All of our parsing, path extraction, weighting scheme, and clustering modules, along with our final results can be found in our replication package, which we provide for verification, reusability and further extension. Weightings can be modified to make the embedding focus on a specific set of elements for building different code embeddings.

>>>>

In this work, we used Ruxanne to successfully mine 20 cross-project bug fix patterns. We presented these patterns in two groups: 12 general patterns and 8 BC-related patterns. Clippy is able to detect five of the BC-related patterns. Fundamentally, Clippy cannot detect some of the remaining patterns. For instance, we proposed the pattern \textit{Adding mutability}. In the Rust context, this change anticipates a future change to the code which will perform a mutation of a previously immutable value. Predicting the future is beyond Clippy's remit. To the best of our knowledge, the rest of the patterns (3 BC-related patterns and 12 general patterns) cannot be reported by the current linting tools.

In each group, we categorized the bug fix patterns based on their underlying program element. The patterns encompass a wide variety of program elements both in the general and BC-related groups. This supports a conclusion that our weighting scheme is not biased towards specific program elements. Also, our patterns are all cross-project fix patterns: each reported fix pattern has been seen at least in three different projects.

@@

All of our parsing, path extraction, weighting scheme, and clustering modules, along with our final results, can be found in our replication package, which we provide for verification, reusability and further extension. Weightings can be modified to make the embedding focus on a specific set of elements for building different code embeddings.

### 2022-11-28 results.tex
minor edit

As discussed in the previous section, we used an embedding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm. We used heuristics (a weighting scheme) while creating our datapoints to make the program elements representing the most important non terminals stand out from the complete set of non terminals seen in the AST. In addition to that, we also incorporated the BC-related elements in our datapoints. Here we evaluate the effectiveness of our embedding.

Our evaluation aims to determine whether the most important program elements get high values in our embedding, i.e. are recognized as important. A human analyzer compares the actual change in the code and compares it to the visual representation of the respective data point. We used circle pack figures to visualize each data point~\citep{collins2003circle}, generated with the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}. In these figures, the radius of the circle for each non terminal shows its importance in our embedding (Figure~\ref{fig:essence} presented an example of the circle pack figures).

We collected random samples 1000 times. In each iteration, we sampled 50 datapoints from our general and borrow-checker databases (25 each). For each database, out of the 1000 sample sets, we picked the one that minimized the sampling error. The sampling error is defined as the average of the absolute differences between the sample set mean and the total mean per column~\citep{degroot2012probability}. Also, these random samples are drawn from the clustered datapoints. The reason behind this decision is the same as why we collected a lot of noise points, that is, the majority of the commits involve changes in many program elements and the key changing elements are difficult to determine.

We carried out a manual experiment for evaluating our code embedding. In the experiment, the human analyzer looks at the change's source code and finds the most important element that appears in the circle pack figure, recording its rank. A value of 1 indicates that the human analyzer concurs with the algorithm, while 2 means that the human-identified most important element was ranked 2 by the algorithm. After collecting all the ranks, then we calculate Top-$n$ Accuracy for $n=1$ up to $n=5$. Both authors acted as analyzer and carried out the experiment independently. To avoid inflating our final results, after comparing the ranks that the authors gave to each datapoint, we consider the larger (worse) rank in our final results. That is, if the first author gave rank 2 to datapoint $D$ and the second author gave rank 3 to that datapoint, in the final results we gave datapoint $D$ a rank of 3. The authors' answers were the same for 45 of the 50 datapoints. The final results are shown in Figure~\ref{fig:rq1}.

@@

\end{axis}
\end{tikzpicture}   
\caption{\label{fig:rq1} Top-$n$ Accuracy of the proposed code embedding approach. The plot shows Top-$n$ Accuracy where $n\in\{1,2,3,4,5\}$ for both general and borrow-checker related datapoints. As shown in the plot, in more than 85\% of the datapoints, the most important element is seen at least in the first three ranks.}

@@

Although we found that our results were acceptable---the result reported by our embedding was top-4 as ranked by humans 88\% of the time---we investigated the other 12\% as well. We found that our embedding tended to fail in change patterns where the change involves many elements (e.g. adding a whole new \verb+impl+ block and implementing new functions in it). It is worth mentioning that these change patterns tend to apply across projects less often, as having many changed elements make it easier for datapoints to be dissimlar.

>>>>

As discussed in the previous section, we used an embedding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm. We used heuristics (in the form of a weighting scheme) while creating our datapoints to make the program elements representing the most important non terminals stand out from the complete set of non terminals seen in the AST. In addition to that, we also incorporated the BC-related elements in our datapoints. Here we evaluate the effectiveness of our embedding.

Our evaluation aims to determine whether the most important program elements get high values in our embedding, i.e. are recognized as important. A human analyzer (paper author) compares the actual change in the code and compares it to the visual representation of the respective data point. We used circle pack figures to visualize each data point~\citep{collins2003circle}, generated with the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}. In these figures, the radius of the circle for each non terminal shows its importance in our embedding (Figure~\ref{fig:essence} presented an example of the circle pack figures).

We collected random samples 1000 times. In each iteration, we sampled 50 datapoints from our general and borrow-checker databases (25 each). For each database, out of the 1000 sample sets, we picked the one that minimized the sampling error (i.e. we are aiming to choose the sample that looks most like the entire dataset). The sampling error is defined as the average of the absolute differences between the sample set mean and the total mean per column~\citep{degroot2012probability}. Also, these random samples are drawn from the clustered datapoints. The reason behind this decision is the same as why we collected a lot of noise points: the majority of the commits involve changes in many program elements and the key changed elements are difficult to determine.

We carried out a manual experiment for evaluating our code embedding. In the experiment, the human analyzer looks at the change's source code and finds the most important element that appears in the circle pack figure, recording its rank. A value of 1 indicates that the human analyzer concurs with the algorithm, while 2 means that the human-identified most important element was ranked 2 by the algorithm. After collecting all the ranks, then we calculate Top-$n$ Accuracy for $n=1$ up to $n=5$. Both authors acted as analyzer and carried out the experiment independently. To avoid inflating our final results, after comparing the ranks that the authors gave to each datapoint, we consider the larger (worse) rank in our final results. That is, if the first author gave rank 2 to datapoint $D$ and the second author gave rank 3 to that datapoint, in the final results we gave datapoint $D$ a rank of 3. The authors' answers were the same for 45 of the 50 datapoints.

@@

\end{axis}
\end{tikzpicture}   
\caption{\label{fig:rq1} Top-$n$ Accuracy of our proposed code embedding approach. The plot shows Top-$n$ Accuracy where $n\in\{1,2,3,4,5\}$ for both general and borrow-checker related datapoints. As shown in the plot, in more than 85\% of the datapoints, the most important element is seen at least in the first three ranks, rising to about 95\% for the first five ranks.}

@@

Figure~\ref{fig:rq1} shows the final results. Although we found that our results were acceptable---the result reported by our embedding was top-4 as ranked by humans 88\% of the time---we investigated the other 12\% as well. We found that our embedding tended to fail in change patterns where the change involves many elements (e.g. adding a whole new \verb+impl+ block and implementing new functions in it). It is worth mentioning that these change patterns tend to apply across projects less often, as having many changed elements make it easier for datapoints to be dissimlar.
